---
title: "data_wrangling"
author: "Kevin Lazenby"
date: "7/20/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
  rmdformats::material:
    highlight: tango
---

This is a script to perform data wrangling on the SRTR data set.
The "function_definitions.Rmd" script must be run prior to running this one.

# Setup

```{r setup, include=FALSE}

knitr::opts_chunk$set(comment = NA, prompt = FALSE, cache = FALSE, echo = TRUE,
                      results = "asis")

library(knitr)
library(tidyverse)
library(haven)
library(rmdformats)
library(dplyr)
library(tidyr)
library(ggpubr)
library(survival)
library(survminer)
library(vistime)
library(stargazer)
library(kableExtra)
library(summarytools)
library(here)

here::i_am("data_wrangling.Rmd")
```

# 1. Importing data (TX_HR, STATJUST_HR1A, STATJUST_HR1B)

```{r import data}

tx_list <- import_tx_data(sub_folder = "pubsaf2103", 
                          file_name = "tx_hr.sas7bdat")

```

# 2. Defining study population

In this section, we select adult, heart-only transplant recipients.

The following exclusion criteria were applied:
1.  Age $\leq$ 17 years old at the time of listing
2.  Recipient of multi-organ transplant

```{r define study population}

tx_list <- apply_exclusion_criteria(input_tx_list = tx_list)

```

# 3. Using SSA Death Master File data to supplement UNOS death records

```{r amend death dates}

tx_list <- supplement_death_records(input_tx_list = tx_list)

```

# 4. Cohort Stratification (Pre-policy cohort #1)

Pre-policy cohort #1:
  - Start date: November 1, 2016
  - End date: October 31, 2017

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r cohort stratification 1}

tx_list_cohort_1 <- do_data_wrangling(input_tx_list = tx_list, 
                  pre_policy_start = as.Date("2016-11-01"),
                  pre_policy_end = as.Date("2017-10-31"),
                  post_policy_start = as.Date("2018-11-01"),
                  post_policy_end = as.Date("2019-10-31"))

(num_tx_per_cohort_1 <- tx_list_cohort_1 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 5. Cohort Stratification (Pre-policy cohort #2)

Pre-policy cohort #2:
  - Start date: November 1, 2017
  - End date: October 17, 2018

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r cohort stratification 2}

tx_list_cohort_2 <- do_data_wrangling(input_tx_list = tx_list, 
                  pre_policy_start = as.Date("2017-11-01"),
                  pre_policy_end = as.Date("2018-10-17"),
                  post_policy_start = as.Date("2018-11-01"),
                  post_policy_end = as.Date("2019-10-31"))

(num_tx_per_cohort_2 <- tx_list_cohort_2 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 6. Cohort Stratification (Pre-policy cohort #3)

Pre-policy cohort #3:
  - Start date: November 1, 2015
  - End date: October 31, 2016
  
Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r cohort stratification 3}

tx_list_cohort_3 <- do_data_wrangling(input_tx_list = tx_list, 
                  pre_policy_start = as.Date("2015-11-01"),
                  pre_policy_end = as.Date("2016-10-31"),
                  post_policy_start = as.Date("2018-11-01"),
                  post_policy_end = as.Date("2019-10-31"))

(num_tx_per_cohort_3 <- tx_list_cohort_3 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 7. Cohort Stratification (Pre-policy cohort #4)

Pre-policy cohort #4:
  - Start date: November 1, 2015
  - End date: October 17, 2018

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r cohort stratification 4}

tx_list_cohort_4 <- do_data_wrangling(input_tx_list = tx_list, 
                  pre_policy_start = as.Date("2015-11-01"),
                  pre_policy_end = as.Date("2018-10-17"),
                  post_policy_start = as.Date("2018-11-01"),
                  post_policy_end = as.Date("2019-10-31"))

(num_tx_per_cohort_4 <- tx_list_cohort_4 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```
