---
title: "kaplan_meier_model"
author: "Kevin Lazenby"
date: "7/21/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
  rmdformats::material:
    highlight: tango
---

This is a script to perform Kaplan-Meier survival modeling.
The script "data_wrangling.Rmd" must be run prior to running this script.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(knitr)
library(tidyverse)
library(haven)
library(rmdformats)
library(dplyr)
library(tidyr)
library(ggpubr)
library(survival)
library(survminer)
library(vistime)
library(stargazer)
library(kableExtra)
library(summarytools)
library(here)
```

# 1. Kaplan-Meier survival model #1

Pre-policy cohort #1:
  - Start date: November 1, 2016
  - End date: October 31, 2017

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r k-m model 1}

post_tx_survival_cohort_1 <- construct_km_model(
  input_cohort_list = tx_list_cohort_1,
  y_limits = c(0.85, 1),
  legend_position = c(0.8, 0.8),
  plot_title = NULL)

post_tx_survival_cohort_1[[2]]

post_tx_survival_cohort_1[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_cohort_1[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)

```

# 2. Kaplan-Meier survival model #2

Pre-policy cohort #2:
  - Start date: November 1, 2017
  - End date: October 17, 2018

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r k-m model 2}

post_tx_survival_cohort_2 <- construct_km_model(
  input_cohort_list = tx_list_cohort_2,
  y_limits = c(0.85, 1),
  legend_position = c(0.9, 0.8),
  plot_title = "1-year K-M survival analysis (pre-policy cohort: 2017-2018)")

post_tx_survival_cohort_2[[2]]

post_tx_survival_cohort_2[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_cohort_2[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 3. Kaplan-Meier survival model #3

Pre-policy cohort #3:
  - Start date: November 1, 2015
  - End date: October 31, 2016

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r k-m model 3}

post_tx_survival_cohort_3 <- construct_km_model(
  input_cohort_list = tx_list_cohort_3,
  y_limits = c(0.85, 1),
  legend_position = c(0.9, 0.8),  
  plot_title = "1-year K-M survival analysis (pre-policy cohort: 2015-2016)")

post_tx_survival_cohort_3[[2]]

post_tx_survival_cohort_3[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_cohort_3[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 4. Kaplan-Meier survival model #4

Pre-policy cohort #4:
  - Start date: November 1, 2015
  - End date: October 17, 2018

Post-policy cohort:
  - Start date: November 1, 2018
  - End date: October 31, 2019

```{r k-m model 4}

post_tx_survival_cohort_4 <- construct_km_model(
  input_cohort_list = tx_list_cohort_4,
  y_limits = c(0.85, 1),
  legend_position = c(0.9, 0.8),  
  plot_title = "1-year K-M survival analysis (pre-policy cohort: 2015-2018)")

post_tx_survival_cohort_4[[2]]

post_tx_survival_cohort_4[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_cohort_4[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```
