---
title: "survival_by_treatment_type"
author: "Kevin Lazenby"
date: "7/20/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
  rmdformats::material:
    highlight: tango
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(knitr)
library(tidyverse)
library(haven)
library(rmdformats)
library(dplyr)
library(tidyr)
library(ggpubr)
library(survival)
library(survminer)
library(vistime)
library(stargazer)
library(kableExtra)
library(summarytools)
library(here)
```

# 1. Kaplan-Meier Survival Model, restricted to recipients with IABP

```{r}

post_tx_survival_iabp <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(IABP == 1),
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_iabp[[2]]

post_tx_survival_iabp[[4]] %>% 
  filter(time >= 363 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_iabp[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 2. Kaplan-Meier Survival Model, restricted to recipients on ECMO

```{r}

post_tx_survival_ecmo <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(ECMO == 1),
  y_limits = c(0.5, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_ecmo[[2]]

post_tx_survival_ecmo[[4]] %>% 
  filter(time >= 265 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_ecmo[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 3. Kaplan-Meier Survival Model, restricted to recipients on low dose inotropes

```{r}

post_tx_survival_low_dose_inotropes <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(low_dose_inotropes == 1),
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_low_dose_inotropes[[2]]

post_tx_survival_low_dose_inotropes[[4]] %>% 
  filter(time >= 363 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_low_dose_inotropes[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```
# 4. Kaplan-Meier Survival Model, restricted to recipients on high dose inotropes

```{r}

post_tx_survival_high_dose_inotropes <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(high_dose_inotropes == 1),
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_high_dose_inotropes[[2]]

post_tx_survival_high_dose_inotropes[[4]] %>% 
  filter(time >= 363 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_high_dose_inotropes[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```


# 5. Kaplan-Meier Survival Model, restricted to recipients on pretransplant mechanical ventilatory support

```{r}

post_tx_survival_ventilator <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(ventilator == 1),
  y_limits = c(0.5, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_ventilator[[2]]

post_tx_survival_ventilator[[4]] %>% 
  filter(time >= 150 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_ventilator[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 6. Kaplan-Meier Survival Model, restricted to recipients with durable LVAD

```{r}

post_tx_survival_durable_lvad <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(durable_LVAD == 1),
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_durable_lvad[[2]]

post_tx_survival_durable_lvad[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_durable_lvad[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 7. Kaplan-Meier Survival Model, restricted to recipients on other MCS

```{r}

post_tx_survival_other_MCS <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(other_MCS == 1),
  y_limits = c(0.75, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_other_MCS[[2]]

post_tx_survival_other_MCS[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_other_MCS[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 8. Kaplan-Meier Survival Model, restricted to recipients on no MCS

```{r}

post_tx_survival_no_MCS <- construct_km_model(
  input_cohort_list = tx_list_cohort_4 %>%
    filter(no_MCS == 1),
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),  
  plot_title = NULL)

post_tx_survival_no_MCS[[2]]

post_tx_survival_no_MCS[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_no_MCS[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)
```

# 9. Assembling Figure 2

```{r}
plot_1A <- post_tx_survival_ecmo[[2]]$plot + 
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_1A <- post_tx_survival_ecmo[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid")
  
plot_1B <- post_tx_survival_iabp[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_1B <- post_tx_survival_iabp[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid")  

plot_1C <- post_tx_survival_ventilator[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") + 
  rremove("legend") +
  rremove("x.grid")

table_1C <- post_tx_survival_ventilator[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") 

plot_1D <- post_tx_survival_durable_lvad[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +
  rremove("legend") +
  rremove("x.grid")

table_1D <- post_tx_survival_durable_lvad[[2]]$table +
  rremove("x.title") +  
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") 

figure_2 <- ggarrange(plot_1A, plot_1B, table_1A, table_1B,
          plot_1C, plot_1D, table_1C, table_1D,
          ncol = 2, nrow = 4, common.legend = TRUE,
          labels = list("A", "B", "", "", "C", "D", "", ""),
          label.x = c(0.05, 0.06, 0, 0, 0.05, 0.06, 0, 0),
          label.y = c(0.275, 0.275, 0, 0, 0.275, 0.275, 0, 0),
          heights = c(0.375, 0.125, 0.375, 0.125, 0.375, 0.125, 0.375, 0.125),
          font.label = list(size = 24, face = "bold")) %>%
  annotate_figure(bottom = text_grob("Time (days)"),
                  left = text_grob("Survival probability", rot = 90))

figure_2

```

# 10. Assembling eFigure 4

```{r}
plot_2A <- post_tx_survival_low_dose_inotropes[[2]]$plot + 
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_2A <- post_tx_survival_low_dose_inotropes[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid")
  
plot_2B <- post_tx_survival_high_dose_inotropes[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_2B <- post_tx_survival_high_dose_inotropes[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid")  

plot_2C <- post_tx_survival_other_MCS[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") + 
  rremove("legend") +
  rremove("x.grid")

table_2C <- post_tx_survival_other_MCS[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") 

plot_2D <- post_tx_survival_no_MCS[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +
  rremove("legend") +
  rremove("x.grid")

table_2D <- post_tx_survival_no_MCS[[2]]$table +
  rremove("x.title") +  
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") 

e_figure <- ggarrange(plot_2A, plot_2B, table_2A, table_2B,
          plot_2C, plot_2D, table_2C, table_2D,
          ncol = 2, nrow = 4, common.legend = TRUE,
          labels = list("A", "B", "", "", "C", "D", "", ""),
          label.x = c(0.06, 0.06, 0, 0, 0.06, 0.06, 0, 0),
          label.y = c(0.275, 0.275, 0, 0, 0.275, 0.275, 0, 0),
          heights = c(0.375, 0.125, 0.375, 0.125, 0.375, 0.125, 0.375, 0.125),
          font.label = list(size = 24, face = "bold")) %>%
  annotate_figure(bottom = text_grob("Time (days)"),
                  left = text_grob("Survival probability", rot = 90))

e_figure

```
