---
title: "truncated_follow_up_demonstration"
author: "Kevin Lazenby"
date: "7/27/2021"
output:
  html_document:
    theme: cosmo
    toc: yes
  rmdformats::material:
    highlight: tango
---

This is a script to demonstrate the effect of truncated follow-up on survival modeling.
The "function_definitions.Rmd" script must be run prior to running this one.

# Setup

```{r setup, include=FALSE}

knitr::opts_chunk$set(comment = NA, prompt = FALSE, cache = FALSE, echo = TRUE,
                      results = "asis")

library(knitr)
library(tidyverse)
library(haven)
library(rmdformats)
library(dplyr)
library(tidyr)
library(ggpubr)
library(survival)
library(survminer)
library(vistime)
library(stargazer)
library(kableExtra)
library(summarytools)
library(here)

here::i_am("truncated_follow_up_demonstration.Rmd")
```

# 1. Importing data (TXF_HR)

In data_wrangling.Rmd, the following operations have already been performed:
- TX_HR was imported
- Study population was defined
- Death records were supplemented
- Cohorts were stratified

```{r import data}

txf_list <- import_txf_data(sub_folder = "pubsaf2103", 
                            file_name = "txf_hr.sas7bdat")

```

# 2. Truncation of follow-up (Post-policy cohort with truncated follow-up #1)
  
```{r truncated cohort 1}

tx_list_truncated_cohort_1 <- truncate_follow_up(
  input_cohort_list = tx_list_cohort_1,
  input_follow_up_list = txf_list,
  analysis_date = as.Date("2019-11-01"))

(num_tx_per_cohort_truncated_1 <- tx_list_truncated_cohort_1 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 3. Truncation of follow-up (Post-policy cohort with truncated follow-up #2)
  
```{r truncated cohort 2}

tx_list_truncated_cohort_2 <- truncate_follow_up(
  input_cohort_list = tx_list_cohort_1,
  input_follow_up_list = txf_list,
  analysis_date = as.Date("2020-05-01"))

(num_tx_per_cohort_truncated_2 <- tx_list_truncated_cohort_2 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 4. Truncation of follow-up (Post-policy cohort with truncated follow-up #3)
  
```{r truncated cohort 3}

tx_list_truncated_cohort_3 <- truncate_follow_up(
  input_cohort_list = tx_list_cohort_1,
  input_follow_up_list = txf_list,
  analysis_date = as.Date("2020-11-01"))

(num_tx_per_cohort_truncated_3 <- tx_list_truncated_cohort_3 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 5. Truncation of follow-up (Post-policy cohort with truncated follow-up #4)
  
```{r truncated cohort 4}

tx_list_truncated_cohort_4 <- truncate_follow_up(
  input_cohort_list = tx_list_cohort_1,
  input_follow_up_list = txf_list,
  analysis_date = as.Date("2021-03-02"))

(num_tx_per_cohort_truncated_4 <- tx_list_truncated_cohort_4 %>% 
  select(PX_ID, policy) %>% 
  group_by(policy) %>% 
  count())
```

# 6. Kaplan-Meier survival model with truncated follow-up #1

```{r k-m model truncated follow-up 1}

post_tx_survival_truncated_cohort_1 <- construct_km_model(
  input_cohort_list = tx_list_truncated_cohort_1,
  y_limits = c(0.7, 1),
  legend_position = c(0.1, 0.2),
  plot_title = NULL)

post_tx_survival_truncated_cohort_1[[2]]

post_tx_survival_truncated_cohort_1[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_truncated_cohort_1[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)

cox_ph_cohort_list <- tx_list_truncated_cohort_1

# run an unadjusted Cox proportional hazards model
cox_ph_model_truncated_1 <- coxph(Surv(surv_time_365, status_365) ~ 
                               post_policy,
                             data = cox_ph_cohort_list)
   
cox_ph_table_truncated_1 <- cox_ph_model_truncated_1 %>% tabcoxph(
  columns = c('beta.se', 'hr.ci', 'z', 'p'),
  var.labels = list(post_policy = "Transplant after policy change"))

cox_ph_model_summary_truncated_1 <- cox_ph_model_truncated_1 %>% summary()

```

# 7. Kaplan-Meier survival model with truncated follow-up #2

```{r k-m model truncated follow-up 2}

post_tx_survival_truncated_cohort_2 <- construct_km_model(
  input_cohort_list = tx_list_truncated_cohort_2,
  y_limits = c(0.85, 1),
  legend_position = c(0.9, 0.8),  
  plot_title = NULL)

post_tx_survival_truncated_cohort_2[[2]]

post_tx_survival_truncated_cohort_2[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_truncated_cohort_2[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)

cox_ph_cohort_list <- tx_list_truncated_cohort_2

# run an unadjusted Cox proportional hazards model
cox_ph_model_truncated_2 <- coxph(Surv(surv_time_365, status_365) ~ 
                               post_policy,
                             data = cox_ph_cohort_list)
   
cox_ph_table_truncated_2 <- cox_ph_model_truncated_2 %>% tabcoxph(
  columns = c('beta.se', 'hr.ci', 'z', 'p'),
  var.labels = list(post_policy = "Transplant after policy change"))

cox_ph_model_summary_truncated_2 <- cox_ph_model_truncated_2 %>% summary()
```

# 8. Kaplan-Meier survival model with truncated follow-up #3

```{r k-m model truncated follow-up 3}

post_tx_survival_truncated_cohort_3 <- construct_km_model(
  input_cohort_list = tx_list_truncated_cohort_3,
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),
  plot_title = NULL)

post_tx_survival_truncated_cohort_3[[2]]

post_tx_survival_truncated_cohort_3[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_truncated_cohort_3[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)

cox_ph_cohort_list <- tx_list_truncated_cohort_3

# run an unadjusted Cox proportional hazards model
cox_ph_model_truncated_3 <- coxph(Surv(surv_time_365, status_365) ~ 
                               post_policy,
                             data = cox_ph_cohort_list)
   
cox_ph_table_truncated_3 <- cox_ph_model_truncated_3 %>% tabcoxph(
  columns = c('beta.se', 'hr.ci', 'z', 'p'),
  var.labels = list(post_policy = "Transplant after policy change"))

cox_ph_model_summary_truncated_3 <- cox_ph_model_truncated_3 %>% summary()

```

# 9. Kaplan-Meier survival model with truncated follow-up #4

```{r k-m model truncated follow-up 4}

post_tx_survival_truncated_cohort_4 <- construct_km_model(
  input_cohort_list = tx_list_truncated_cohort_4,
  y_limits = c(0.85, 1),
  legend_position = c(0.1, 0.2),
  plot_title = NULL)

post_tx_survival_truncated_cohort_4[[2]]

post_tx_survival_truncated_cohort_4[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 0) %>% 
  select(post_policy, time, lower, surv, upper)

post_tx_survival_truncated_cohort_4[[4]] %>% 
  filter(time >= 364 & time <= 365, post_policy == 1) %>% 
  select(post_policy, time, lower, surv, upper)

cox_ph_cohort_list <- tx_list_truncated_cohort_4

# run an unadjusted Cox proportional hazards model
cox_ph_model_truncated_4 <- coxph(Surv(surv_time_365, status_365) ~ 
                               post_policy,
                             data = cox_ph_cohort_list)
   
cox_ph_table_truncated_4 <- cox_ph_model_truncated_4 %>% tabcoxph(
  columns = c('beta.se', 'hr.ci', 'z', 'p'),
  var.labels = list(post_policy = "Transplant after policy change"))

cox_ph_model_summary_truncated_4 <- cox_ph_model_truncated_4 %>% summary()
```

# 10. Assembling Figure 2

```{r}
plot_e1A <- post_tx_survival_truncated_cohort_1[[2]]$plot + 
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_e1A <- post_tx_survival_truncated_cohort_1[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") +
  theme(axis.text = element_text(size = 24), title = element_text(size = 24))
  
plot_e1B <- post_tx_survival_truncated_cohort_2[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +  
  rremove("legend") +
  rremove("x.grid")

table_e1B <- post_tx_survival_truncated_cohort_2[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") +
  theme(axis.text = element_text(size = 24), title = element_text(size = 24))

plot_e1C <- post_tx_survival_truncated_cohort_3[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") + 
  rremove("legend") +
  rremove("x.grid")

table_e1C <- post_tx_survival_truncated_cohort_3[[2]]$table +
  rremove("x.title") +
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") +
  theme(axis.text = element_text(size = 24), title = element_text(size = 24))

plot_e1D <- post_tx_survival_truncated_cohort_4[[2]]$plot +
  rremove("x.title") +
  rremove("y.title") +
  rremove("legend") +
  rremove("x.grid")

table_e1D <- post_tx_survival_truncated_cohort_4[[2]]$table +
  rremove("x.title") +  
  rremove("y.title") +
  rremove("x.grid") +
  rremove("y.grid") +
  theme(axis.text = element_text(size = 24), title = element_text(size = 24))

figure_2 <- ggarrange(plot_e1A, plot_e1B, table_e1A, table_e1B,
          plot_e1C, plot_e1D, table_e1C, table_e1D,
          ncol = 2, nrow = 4, common.legend = TRUE,
          labels = list("A", "B", "", "", "C", "D", "", ""),
          label.x = c(0.05, 0.06, 0, 0, 0.05, 0.06, 0, 0),
          label.y = c(0.3, 0.3, 0, 0, 0.3, 0.3, 0, 0),
          heights = c(0.375, 0.125, 0.375, 0.125, 0.375, 0.125, 0.375, 0.125),
          font.label = list(size = 96, face = "bold")) %>%
  annotate_figure(bottom = text_grob("Time (days)", face = "bold", size = 30),
                  left = text_grob("Survival probability", rot = 90, face = "bold", size = 30))

```
